---
title: "Storyboard Jari Kamps"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
---

Homework Week 11 {.storyboard}
=========================================

### Tempograms
For this weeks homework I made two tempograms based on two outliers from the corpus. I compared the tracks with the highest and lowest bpm and made both tempograms. The track from the corpus with the lowest bpm is Mileage with a 61.778 bpm, the highest is Middle with 199.928 bpm. PS I have some major technical difficulties with the loading and knitting of my tempograms. 

```{r}
```

### Mileage

```{r}
library(tidyverse)
library(spotifyr)
library(compmus)

Mileage <-
  get_tidy_audio_analysis("4tOoKxtiLr1qH2AHUI5KqU") |>
  select(segments) |>
  unnest(segments)

Mileage |>
  mutate(loudness_max_time = start + loudness_max_time) |>
  arrange(loudness_max_time) |>
  mutate(delta_loudness = loudness_max - lag(loudness_max)) |>
  ggplot(aes(x = loudness_max_time, y = pmax(0, delta_loudness))) +
  geom_line() +
  xlim(0, 30) +
  theme_minimal() +
  labs(x = "Time (s)", y = "Novelty")

```

### Middle

```{r}
library(tidyverse)
library(spotifyr)
library(compmus)

Middle <-
  get_tidy_audio_analysis("0JFkkqgv6r2SpbEAj7q5rt") |>
  select(segments) |>
  unnest(segments)

Middle |>
  mutate(loudness_max_time = start + loudness_max_time) |>
  arrange(loudness_max_time) |>
  mutate(delta_loudness = loudness_max - lag(loudness_max)) |>
  ggplot(aes(x = loudness_max_time, y = pmax(0, delta_loudness))) +
  geom_line() +
  xlim(0, 30) +
  theme_minimal() +
  labs(x = "Time (s)", y = "Novelty")

```

Introduction {.storyboard}
=========================================

### Introduction to corpus

```{r}
```

### Frame 2
Hier kan je je tekst zetten.

```{r}
```

Graphs {.storyboard}
=========================================

### Frame 3

```{r}
```

### Frame 4

```{r}
```

Column
-----------------------------------------