---
title: "The commercial success of solo albums by Sevn Alias. Made by Jari Kamps"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
---

Introduction {.storyboard}
=========================================

### INTRODUCTION TO THE CORPUS
First of all, a description of my corpus to start with my portfolio for computational musicology. I chose for one of the biggest Dutch Hip Hop artists of the past decade, Sevn Alias, as the main focus of my corpus, with the related question: ‘’Which Sevn Alias albums were the most commercially successful and does this commercial success correspond in any way to the label where the artist was at the time signed at?’’. I chose this corpus of all his Spotify available solo-albums because Sevn Alias was one of the first Dutch hip hop artists who started popping up at the beginning of the rise of the influence of hip hop in the Dutch music scene, and he is considered to be one of the biggest and most influential (based on his music and the cultural influence around it). The main interesting part of it is, in my opinion, the fact that he released over 10 albums in the past decade, all with major success and most of them distributed under different labels in each of the release periods. 
The natural groups and points of comparison in the corpus are mainly focussed on the artist Sevn Alias his solo albums, the music labels who distributed the albums and the amounts of streams of each track from every album. I think these groups are most relatable on the point that it is the same artist on every track and differ the most on the producers behind the tracks and possible featurings on the tracks. Although, these can also be the main points that I’m still a bit uncertain about, because they also have an influence on the outcome of each track and therefore, the possible commercial success. 
I created my own Spotify-playlist with all available tracks from every solo-album of Sevn Alias available on solely Spotify. The following link leads to the public playlist I created: https://open.spotify.com/playlist/36UKFGHMewsc9n8KIspNrB?si=bdfe0cdf208347fb

In relation to the structure of the corpus, there are a couple of sidenotes i have to adress. First of all, his very first solo album 'TwentyFourSevn' is not available on Spotify, so this won't be applicable in the research. Secondly, the track called 'Intro' from the album 'Trap Grammy' is also not applicable, since it gave an error rendering some of the graphs related to the corpus since there is one more track with exactly the same track name. Thirdly, I had some issues with the coding of making my graphs and grams 'interactual', so it could be that these functions don't work unfortunately.

Last of all, a quick overview of the albums and the labels who distributed them, in order of release date (2016 - 2022).

Twenty Four Sevn 2 - Rotterdam Airlines

Twenty Four Sevn 3 - Rotterdam Airlines, Cloud 9

Trap Grammy - Rotterdam Airlines, Sony Music

Picasso - Rotterdam Airlines

OAKINN - 1OAK Entertainment

Recasso - 1OAK Entertainment 

SIRIUS - 1OAK Entertainment 

Twenty Four Sevn 4 - 1OAK Entertainment 

Bogey 4 the Win - 1OAK Entertainment 

Six Summers - 1OAK Entertainment 

The Last Flowbender - 1OAK Entertainment 

'At the least half of the '1OAK Entertainment' releases in joint venture with Top Notch Music, Universal and the most recent half fully independent as his own label'

```{r}
```

First impressions {.storyboard}
=========================================

### FIRST IMPRESSIONS
For my first Spotify API impressions of the corpus, I chose to create several ggplots, since I think that three track features are the support for my topic. In this case the danceability, energy and track popularity API features. I think the track features danceability and energy could play an important factor in terms of commercial success, because I think these factors are important for things like the extent to which it is played in public spheres. A higher danceability and energy level could perhaps support the definition of commercial success per album. The track popularity (number between 0 and 100) represents the overall popularity of the song by album, with a number closer to 100 being Sevn Alias most popular song. 

By first impression, we can see that the album 'Trap Grammy' is one of the least popular albums and the most recent album 'The Last Flowbender' is one of the most popular. If we relate this to both albums general danceability and energy level, we see some interesting aspects. The overal energy and danceability levels of both albums seems to be at the same level, but the album 'Trap Grammy' has a wider spread of tracks with some higher energy and danceabilty. This seems to be relateable that the listeners of Sevn Alias' music relate the track popularity more to tracks with less danceability. Maybe they relate more to a specific style of Rap in the music, than music with high energy and they can dance to.

if we relate the track popularity gram to the albums of the corpus based on streams, we see that it has little relation to the amount of streams an album has to this point in time. For example the album 'Picasso' has songs with over 56 million streams and a minimum of 2.5, but the latest album 'The Last Flowbender' a track over 1.4 and a minimun of 198.000. Eventhough it shows in the gram that the median track populariy is a lot higher in the latest solo album.

I also find it interesting that the energy and danceability levels are somewhat different in comparison, especially in the releases between 2018 and 2019, where a higher degree of danceability corelates to a lower energy. It is also interesting that the general track popularity of the songs has a relatively general point per album, despite huge hits, 'outliers', in for example the year 2018. Around 2021, the album has a broader spread in terms of general popularity. I mainly found these aspects interesting.

```{r}
```

### VISUALISATION TRACK POPULARITY

```{r}
library(tufte)
library(spotifyr)
library(tidyverse)
library(ggplot2)
library(dplyr)

sevn <- get_playlist_audio_features("", "36UKFGHMewsc9n8KIspNrB")

ggplot(sevn, aes(y = track.popularity, x = as.Date(track.album.release_date))) +
  geom_point() +
  geom_smooth() +
  xlim(as.Date("2016-01-15"), as.Date("2022-12-09")) +
  labs(x = "Release Date", y = "Track Popularity", title = "TRACK POPULARITY LEVEL") +
  theme_light()

```

### VISUALISATION ENERGY

```{r}
library(tufte)
library(spotifyr)
library(tidyverse)
library(ggplot2)
library(dplyr)

sevn <- get_playlist_audio_features("", "36UKFGHMewsc9n8KIspNrB")

ggplot(sevn, aes(y = energy, x = as.Date(track.album.release_date))) +
  geom_point() +
  geom_smooth() +
  xlim(as.Date("2016-01-15"), as.Date("2022-12-09")) +
  labs(x = "Release Date", y = "Energy", title = "ENERGY LEVEL") +
  theme_light()

```

### VISUALISATION DANCEABILITY

```{r}
library(tufte)
library(spotifyr)
library(tidyverse)
library(ggplot2)
library(dplyr)

sevn <- get_playlist_audio_features("", "36UKFGHMewsc9n8KIspNrB")

ggplot(sevn, aes(y = danceability, x = as.Date(track.album.release_date))) +
  geom_point() +
  geom_smooth() +
  xlim(as.Date("2016-01-15"), as.Date("2022-12-09")) +
  labs(x = "Release Date", y = "Danceability", title = "DANCEABILITY LEVEL") +
  theme_light()

```

Noticeable outliers {.storyboard}
=========================================

### NOTICEABLE OUTLIERS
Looking forward to making a chromagram visualisation on some ‘outlier’ in the corpus i chose to look at an outlier based on the key of the tracks and the correlating track popularity feature. What struck me when I looked at the data was that in this case there were two main tracks as outliers in the entire corpus. The track ‘Layhoun’ and the track ‘Centen’, both from the album ‘Twenty Four Sevn 3’ are the only two tracks with a specific key that no other track in the entire corpus has. The track ‘Layhoun’ is the only track in the key E major and the track ‘Centen’ is the only track in the key B major. Interesting since almost all tracks in the corpus are based on a minor key in general. Another interesting fact is that both of the tracks in relation to this have a very low track popularity level of just 15 for ‘Layhoun’ and 14 for ‘Centen’ while the rest of the tracks from the same album all have a relative high track popularity.

In the first visualisation is a chromagram of the track ‘Layhoun’ visible. One thing that is really remarkeable is that the chromagram shows a consistent ‘half’ magnitude for the scale of E major, but shows a lot of peak magnitudes for the G#/Ab key. When I re-heard the song myself, I came to the conclusion that the track contains a lot of dissonance and intonation features in the main melody, also in relation to the change of octaves. This may cause the peaks of magnitudes in keys which are not directly related to the main scale of the track itself. 

On the other hand, shows the chromagram of the track ‘Centen’ a very clear magnitude on the scale of the track B major. When I re-heard this track, I remarked that the track sounds quite monotonous throughout the entire track so in the end it makes sense.

```{r}
```

### VISUALISATION CHROMAGRAM LAYHOUN

```{r}
library(tidyverse)
library(spotifyr)
library(compmus)

layhoun <-
  get_tidy_audio_analysis("74ebZ2HIgo4NmTy9HMe0oN") |>
  select(segments) |>
  unnest(segments) |>
  select(start, duration, pitches)

layhoun |>
  mutate(pitches = map(pitches, compmus_normalise, "euclidean")) |>
  compmus_gather_chroma() |> 
  ggplot(
    aes(
      x = start + duration / 2,
      width = duration,
      y = pitch_class,
      fill = value
    )
  ) +
  geom_tile() +
  labs(x = "Time (s)", y = NULL, fill = "Magnitude") +
  theme_minimal() +
  scale_fill_viridis_c()

```

### VISUALISATION CHROMAGRAM CENTEN

```{r}
library(tidyverse)
library(spotifyr)
library(compmus)

centen <-
  get_tidy_audio_analysis("6cc3CUymJMeVrPUPsRjVad") |>
  select(segments) |>
  unnest(segments) |>
  select(start, duration, pitches)

centen |>
  mutate(pitches = map(pitches, compmus_normalise, "euclidean")) |>
  compmus_gather_chroma() |> 
  ggplot(
    aes(
      x = start + duration / 2,
      width = duration,
      y = pitch_class,
      fill = value
    )
  ) +
  geom_tile() +
  labs(x = "Time (s)", y = NULL, fill = "Magnitude") +
  theme_minimal() +
  scale_fill_viridis_c()

```

Timbre outliers {.storyboard}
=========================================

### TIMBRE OUTLIERS
When had a look on working on the ‘timbre’ feature, I decided to base the research on outliers from the corpus based on contribution to this timbre. To me this came down to, among other things, a distinction in the ‘Instrumentalness’ feature. First of all, I made a visualisation to identify what the outliers are applicable for the timbre research. From the graph can be seen that there are clearly two outliers. In this case the track ‘Ik Trap’ from the album OAKINN with the most instrumentalness and secondly the track ‘Klim’ from the album SIRIUS. Based on further data, both tracks have an average danceability of around 0.8, but have a major difference in energy level. ‘Ik Trap’ contains a level of 0.7 and ‘Klim’ a level of 0.4.

Visualisations for ‘Ik Trap’

From the ‘Ik Trap’ cepstogram visualisation is mainly visible that the track plays mainly between the C2 and C5 range of pitch. The magnitude peaks are through the time of the piece mainly focussed on the C4 range in the beginning, but transforms and evaluates more to the range of C5 in the later parts. Interesting about this gram is that in the beginning the C0 plays very little to no role at all, but comes up at around 20 seconds in the pieces. The self-similarity matrix on the other hand confirms that the track makes use of a lot of same-sounding components and repetition from this 20 second point with the dark parallel line and darker boxing. 

Visualisations for ‘Klim’

From the ‘Klim’ cepstogram visualisation is mainly visible that the track plays mainly between the C2 and C4 range of pitch. Interesting is that the piece stays constant between this range through almost the entire piece, with some major peaks of magnitude in the C2 range at around 140 seconds. This consistency also relates to the self-similarity matrix, where its almost dark in all the piece, except for a major break at this 140 second mark, where it is not similar at all. 

When we relate the cepstograms of the two tracks to each other, we see that ‘Ik Trap’ has more flexibility of pitch range throughout the track than ‘Klim’, even though ‘Klim’ is a more consistent piece. When we relate the self-similarity matrices to each other we can conclude around the same things. 

```{r}
```

### VISUALISATION INSTRUMENTALNESS

```{r}
library(tufte)
library(spotifyr)
library(tidyverse)
library(ggplot2)
library(dplyr)

sevn <- get_playlist_audio_features("", "36UKFGHMewsc9n8KIspNrB")

ggplot(sevn, aes(y = instrumentalness, x = as.Date(track.album.release_date))) +
  geom_point() +
  geom_smooth() +
  xlim(as.Date("2016-01-15"), as.Date("2022-12-09")) +
  labs(x = "Release Date", y = "Instrumentalness", title = "INSTRUMENTALNESS LEVEL") +
  theme_light()

```

### VISUALISATION IK TRAP CEPSTOGRAM

```{r}
library(tidyverse)
library(spotifyr)
library(compmus)

iktrap <-
  get_tidy_audio_analysis("0unMc21fkiIASp7FjgWJ9Z") |> # Change URI.
  compmus_align(bars, segments) |>                     # Change `bars`
  select(bars) |>                                      #   in all three
  unnest(bars) |>                                      #   of these lines.
  mutate(
    pitches =
      map(segments,
          compmus_summarise, pitches,
          method = "rms", norm = "euclidean"              # Change summary & norm.
      )
  ) |>
  mutate(
    timbre =
      map(segments,
          compmus_summarise, timbre,
          method = "rms", norm = "euclidean"              # Change summary & norm.
      )
  )

iktrap |>
  compmus_gather_timbre() |>
  ggplot(
    aes(
      x = start + duration / 2,
      width = duration,
      y = basis,
      fill = value
    )
  ) +
  geom_tile() +
  labs(x = "Time (s)", y = NULL, fill = "Magnitude") +
  scale_fill_viridis_c() +                              
  theme_classic()

```

### VISUALISATION IK TRAP SELF-SIMILARITY MATRIX

```{r}
library(tidyverse)
library(spotifyr)
library(compmus)

iktrap <-
  get_tidy_audio_analysis("0unMc21fkiIASp7FjgWJ9Z") |> # Change URI.
  compmus_align(bars, segments) |>                     # Change `bars`
  select(bars) |>                                      #   in all three
  unnest(bars) |>                                      #   of these lines.
  mutate(
    pitches =
      map(segments,
          compmus_summarise, pitches,
          method = "rms", norm = "euclidean"              # Change summary & norm.
      )
  ) |>
  mutate(
    timbre =
      map(segments,
          compmus_summarise, timbre,
          method = "rms", norm = "euclidean"              # Change summary & norm.
      )
  )

iktrap |>
  compmus_self_similarity(timbre, "cosine") |> 
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  scale_fill_viridis_c(guide = "none") +
  theme_classic() +
  labs(x = "", y = "")

```

### VISUALISATION KLIM CEPSTOGRAM

```{r}
library(tidyverse)
library(spotifyr)
library(compmus)

klim <-
  get_tidy_audio_analysis("5WZTHLjL4kz8rdBJzGtVpP") |> # Change URI.
  compmus_align(bars, segments) |>                     # Change `bars`
  select(bars) |>                                      #   in all three
  unnest(bars) |>                                      #   of these lines.
  mutate(
    pitches =
      map(segments,
          compmus_summarise, pitches,
          method = "rms", norm = "euclidean"              # Change summary & norm.
      )
  ) |>
  mutate(
    timbre =
      map(segments,
          compmus_summarise, timbre,
          method = "rms", norm = "euclidean"              # Change summary & norm.
      )
  )

klim |>
  compmus_gather_timbre() |>
  ggplot(
    aes(
      x = start + duration / 2,
      width = duration,
      y = basis,
      fill = value
    )
  ) +
  geom_tile() +
  labs(x = "Time (s)", y = NULL, fill = "Magnitude") +
  scale_fill_viridis_c() +                              
  theme_classic()

```

### VISUALISATION KLIM SELF-SIMILARITY MATRIX

```{r}
library(tidyverse)
library(spotifyr)
library(compmus)

klim <-
  get_tidy_audio_analysis("5WZTHLjL4kz8rdBJzGtVpP") |> # Change URI.
  compmus_align(bars, segments) |>                     # Change `bars`
  select(bars) |>                                      #   in all three
  unnest(bars) |>                                      #   of these lines.
  mutate(
    pitches =
      map(segments,
          compmus_summarise, pitches,
          method = "rms", norm = "euclidean"              # Change summary & norm.
      )
  ) |>
  mutate(
    timbre =
      map(segments,
          compmus_summarise, timbre,
          method = "rms", norm = "euclidean"              # Change summary & norm.
      )
  )

klim |>
  compmus_self_similarity(timbre, "cosine") |> 
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  scale_fill_viridis_c(guide = "none") +
  theme_classic() +
  labs(x = "", y = "")

```

Track-level summaries {.storyboard}
=========================================

### TRACK-LEVEL SUMMARIES
Incorporating more track-level summaries into the research led me to creating several chordograms of two outliers in the corpus, a graph based on tempo, loudness and duration and a graph based on timbre coefficients for the two most recent albums in the corpus. 

First of all, a look at the keygrams. I chose to pick two outliers from the corpus based on the highest and lowest valence level, due to the relation to pitch and key. The track ‘Mileage’ from the album Six Summers has the highest valence level of 0.962 and the track ‘Stappen’ from the album Twenty Four Sevn 2 has the lowest valence level of 0.123. When we take a look at the keygram of ‘Mileage’ we see that the key throughout the entire piece is mainly based on the B minor key, even though the first 10 songs are indicated in the Ab major key, which indicates that there might be a certain intro with a modulation after it from major to minor.  This is since the ‘happiness’ is supposed to be really high, but the minor key suggests a sad feeling from the track. The keygram of the track ‘Stappen’ shows a clear distinction between the major and minor scale, where minor has a clear dominance. 

The visualisation based on tempo, loudness and duration is based on the average most popular albums from the track popularity graph we discussed earlier. This is the case for the two most recent albums ‘The Last Flowbender’ and ‘Six Summers’. The graph indicates that most of the tracks from The Last Flowbender are based on a higher BPM than the tracks from Six Summers. The duration of the tracks are more or less overall the same, but the loudness of the tracks differs significantly. The average loudness of the tracks from the Six Summers album is significantly than the tracks from The Last Flowbender album. This can correlate to the fact that the popularity of track from the Six Summers album is much more pulled together than the wider spread from The Last Flowbender album. 

The timbre coefficient graph shows that both albums are quite similar based on timbre. Even though not every ‘bulb’ is stretched with the same length, they show similarity in the way the stand next to each other and follow ‘the same route’ through the graph.

```{r}
```

### VISUALISATION MILEAGE KEYGRAM

```{r}
library(tidyverse)
library(spotifyr)
library(compmus)

circshift <- function(v, n) {
  if (n == 0) v else c(tail(v, n), head(v, -n))
}

#      C     C#    D     Eb    E     F     F#    G     Ab    A     Bb    B
major_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    0,    0)
minor_chord <-
  c(   1,    0,    0,    1,    0,    0,    0,    1,    0,    0,    0,    0)
seventh_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    1,    0)

major_key <-
  c(6.35, 2.23, 3.48, 2.33, 4.38, 4.09, 2.52, 5.19, 2.39, 3.66, 2.29, 2.88)
minor_key <-
  c(6.33, 2.68, 3.52, 5.38, 2.60, 3.53, 2.54, 4.75, 3.98, 2.69, 3.34, 3.17)

chord_templates <-
  tribble(
    ~name, ~template,
    "Gb:7", circshift(seventh_chord, 6),
    "Gb:maj", circshift(major_chord, 6),
    "Bb:min", circshift(minor_chord, 10),
    "Db:maj", circshift(major_chord, 1),
    "F:min", circshift(minor_chord, 5),
    "Ab:7", circshift(seventh_chord, 8),
    "Ab:maj", circshift(major_chord, 8),
    "C:min", circshift(minor_chord, 0),
    "Eb:7", circshift(seventh_chord, 3),
    "Eb:maj", circshift(major_chord, 3),
    "G:min", circshift(minor_chord, 7),
    "Bb:7", circshift(seventh_chord, 10),
    "Bb:maj", circshift(major_chord, 10),
    "D:min", circshift(minor_chord, 2),
    "F:7", circshift(seventh_chord, 5),
    "F:maj", circshift(major_chord, 5),
    "A:min", circshift(minor_chord, 9),
    "C:7", circshift(seventh_chord, 0),
    "C:maj", circshift(major_chord, 0),
    "E:min", circshift(minor_chord, 4),
    "G:7", circshift(seventh_chord, 7),
    "G:maj", circshift(major_chord, 7),
    "B:min", circshift(minor_chord, 11),
    "D:7", circshift(seventh_chord, 2),
    "D:maj", circshift(major_chord, 2),
    "F#:min", circshift(minor_chord, 6),
    "A:7", circshift(seventh_chord, 9),
    "A:maj", circshift(major_chord, 9),
    "C#:min", circshift(minor_chord, 1),
    "E:7", circshift(seventh_chord, 4),
    "E:maj", circshift(major_chord, 4),
    "G#:min", circshift(minor_chord, 8),
    "B:7", circshift(seventh_chord, 11),
    "B:maj", circshift(major_chord, 11),
    "D#:min", circshift(minor_chord, 3)
  )

key_templates <-
  tribble(
    ~name, ~template,
    "Gb:maj", circshift(major_key, 6),
    "Bb:min", circshift(minor_key, 10),
    "Db:maj", circshift(major_key, 1),
    "F:min", circshift(minor_key, 5),
    "Ab:maj", circshift(major_key, 8),
    "C:min", circshift(minor_key, 0),
    "Eb:maj", circshift(major_key, 3),
    "G:min", circshift(minor_key, 7),
    "Bb:maj", circshift(major_key, 10),
    "D:min", circshift(minor_key, 2),
    "F:maj", circshift(major_key, 5),
    "A:min", circshift(minor_key, 9),
    "C:maj", circshift(major_key, 0),
    "E:min", circshift(minor_key, 4),
    "G:maj", circshift(major_key, 7),
    "B:min", circshift(minor_key, 11),
    "D:maj", circshift(major_key, 2),
    "F#:min", circshift(minor_key, 6),
    "A:maj", circshift(major_key, 9),
    "C#:min", circshift(minor_key, 1),
    "E:maj", circshift(major_key, 4),
    "G#:min", circshift(minor_key, 8),
    "B:maj", circshift(major_key, 11),
    "D#:min", circshift(minor_key, 3)
  )

mileage <-
  get_tidy_audio_analysis("4tOoKxtiLr1qH2AHUI5KqU") |>
  compmus_align(sections, segments) |>
  select(sections) |>
  unnest(sections) |>
  mutate(
    pitches =
      map(segments,
          compmus_summarise, pitches,
          method = "mean", norm = "manhattan"
      )
  )

mileage |> 
  compmus_match_pitch_template(
    key_templates,         # Change to chord_templates if descired
    method = "euclidean",  # Try different distance metrics
    norm = "manhattan"     # Try different norms
  ) |>
  ggplot(
    aes(x = start + duration / 2, width = duration, y = name, fill = d)
  ) +
  geom_tile() +
  scale_fill_viridis_c(guide = "none") +
  theme_minimal() +
  labs(x = "Time (s)", y = "")

```

### VISUALISATION STAPPEN KEYGRAM

```{r}
library(tidyverse)
library(spotifyr)
library(compmus)

circshift <- function(v, n) {
  if (n == 0) v else c(tail(v, n), head(v, -n))
}

#      C     C#    D     Eb    E     F     F#    G     Ab    A     Bb    B
major_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    0,    0)
minor_chord <-
  c(   1,    0,    0,    1,    0,    0,    0,    1,    0,    0,    0,    0)
seventh_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    1,    0)

major_key <-
  c(6.35, 2.23, 3.48, 2.33, 4.38, 4.09, 2.52, 5.19, 2.39, 3.66, 2.29, 2.88)
minor_key <-
  c(6.33, 2.68, 3.52, 5.38, 2.60, 3.53, 2.54, 4.75, 3.98, 2.69, 3.34, 3.17)

chord_templates <-
  tribble(
    ~name, ~template,
    "Gb:7", circshift(seventh_chord, 6),
    "Gb:maj", circshift(major_chord, 6),
    "Bb:min", circshift(minor_chord, 10),
    "Db:maj", circshift(major_chord, 1),
    "F:min", circshift(minor_chord, 5),
    "Ab:7", circshift(seventh_chord, 8),
    "Ab:maj", circshift(major_chord, 8),
    "C:min", circshift(minor_chord, 0),
    "Eb:7", circshift(seventh_chord, 3),
    "Eb:maj", circshift(major_chord, 3),
    "G:min", circshift(minor_chord, 7),
    "Bb:7", circshift(seventh_chord, 10),
    "Bb:maj", circshift(major_chord, 10),
    "D:min", circshift(minor_chord, 2),
    "F:7", circshift(seventh_chord, 5),
    "F:maj", circshift(major_chord, 5),
    "A:min", circshift(minor_chord, 9),
    "C:7", circshift(seventh_chord, 0),
    "C:maj", circshift(major_chord, 0),
    "E:min", circshift(minor_chord, 4),
    "G:7", circshift(seventh_chord, 7),
    "G:maj", circshift(major_chord, 7),
    "B:min", circshift(minor_chord, 11),
    "D:7", circshift(seventh_chord, 2),
    "D:maj", circshift(major_chord, 2),
    "F#:min", circshift(minor_chord, 6),
    "A:7", circshift(seventh_chord, 9),
    "A:maj", circshift(major_chord, 9),
    "C#:min", circshift(minor_chord, 1),
    "E:7", circshift(seventh_chord, 4),
    "E:maj", circshift(major_chord, 4),
    "G#:min", circshift(minor_chord, 8),
    "B:7", circshift(seventh_chord, 11),
    "B:maj", circshift(major_chord, 11),
    "D#:min", circshift(minor_chord, 3)
  )

key_templates <-
  tribble(
    ~name, ~template,
    "Gb:maj", circshift(major_key, 6),
    "Bb:min", circshift(minor_key, 10),
    "Db:maj", circshift(major_key, 1),
    "F:min", circshift(minor_key, 5),
    "Ab:maj", circshift(major_key, 8),
    "C:min", circshift(minor_key, 0),
    "Eb:maj", circshift(major_key, 3),
    "G:min", circshift(minor_key, 7),
    "Bb:maj", circshift(major_key, 10),
    "D:min", circshift(minor_key, 2),
    "F:maj", circshift(major_key, 5),
    "A:min", circshift(minor_key, 9),
    "C:maj", circshift(major_key, 0),
    "E:min", circshift(minor_key, 4),
    "G:maj", circshift(major_key, 7),
    "B:min", circshift(minor_key, 11),
    "D:maj", circshift(major_key, 2),
    "F#:min", circshift(minor_key, 6),
    "A:maj", circshift(major_key, 9),
    "C#:min", circshift(minor_key, 1),
    "E:maj", circshift(major_key, 4),
    "G#:min", circshift(minor_key, 8),
    "B:maj", circshift(major_key, 11),
    "D#:min", circshift(minor_key, 3)
  )

stappen <-
  get_tidy_audio_analysis("0aFlFn9MFDtDLrHVgxQGct") |>
  compmus_align(sections, segments) |>
  select(sections) |>
  unnest(sections) |>
  mutate(
    pitches =
      map(segments,
          compmus_summarise, pitches,
          method = "mean", norm = "manhattan"
      )
  )

stappen |> 
  compmus_match_pitch_template(
    key_templates,         # Change to chord_templates if descired
    method = "euclidean",  # Try different distance metrics
    norm = "manhattan"     # Try different norms
  ) |>
  ggplot(
    aes(x = start + duration / 2, width = duration, y = name, fill = d)
  ) +
  geom_tile() +
  scale_fill_viridis_c(guide = "none") +
  theme_minimal() +
  labs(x = "Time (s)", y = "")

```

### VISUALISATION ALBUMS TEMPO, LOUDNESS, DURATION

```{r}
library(tidyverse)
library(spotifyr)
library(compmus)

thelastflowbender <-
  get_playlist_audio_features(
    "jnk",
    "5qAUrZRUPqnYCRksFe2uXu"
  ) |>
  slice(1:30) |>
  add_audio_analysis()
sixsummers <-
  get_playlist_audio_features(
    "jnk",
    "2rwSVIdnBZN3IxflZSeNGv"
  ) |>
  slice(1:30) |>
  add_audio_analysis()
allalbums <-
  thelastflowbender |>
  mutate(album = "The Last Flowbender") |>
  bind_rows(sixsummers |> mutate(album = "Six Summers"))

allalbums |>
  mutate(
    sections =
      map(
        sections,                                    # sections or segments
        summarise_at,
        vars(tempo, loudness, duration),             # features of interest
        list(section_mean = mean, section_sd = sd)   # aggregation functions
      )
  ) |>
  unnest(sections) |>
  ggplot(
    aes(
      x = tempo,
      y = tempo_section_sd,
      colour = album,
      alpha = loudness
    )
  ) +
  geom_point(aes(size = duration / 60)) +
  geom_rug() +
  theme_minimal() +
  ylim(0, 5) +
  labs(
    x = "Mean Tempo (bpm)",
    y = "SD Tempo",
    colour = "Album",
    size = "Duration (min)",
    alpha = "Volume (dBFS)"
  )

```

### VISUALISATION ALBUMS TIMBRE COEFFICIENTS

```{r}
library(tidyverse)
library(spotifyr)
library(compmus)

thelastflowbender <-
  get_playlist_audio_features(
    "jnk",
    "5qAUrZRUPqnYCRksFe2uXu"
  ) |>
  slice(1:30) |>
  add_audio_analysis()
sixsummers <-
  get_playlist_audio_features(
    "jnk",
    "2rwSVIdnBZN3IxflZSeNGv"
  ) |>
  slice(1:30) |>
  add_audio_analysis()
allalbums <-
  thelastflowbender |>
  mutate(album = "The Last Flowbender") |>
  bind_rows(sixsummers |> mutate(album = "Six Summers"))

allalbums |>
  mutate(
    timbre =
      map(
        segments,
        compmus_summarise,
        timbre,
        method = "mean"
      )
  ) |>
  select(album, timbre) |>
  compmus_gather_timbre() |>
  ggplot(aes(x = basis, y = value, fill = album)) +
  geom_violin() +
  scale_fill_viridis_d() +
  labs(x = "Spotify Timbre Coefficients", y = "", fill = "Album")

```

Tempo and loudness novelty {.storyboard}
=========================================

### TEMPO AND LOUDNESS NOVELTY
I discussed with my TA Dave Baker that my grams/graphs in relation to 'Tempo' are not able to render on my laptop and we couldn't figure out a solution. This has been communicated if I'm not mistaking. Nevertheless, I made two novelty functions based on two tracks from the corpus. I compared the tracks with the highest and lowest BPM and made both novelty functions. The track from the corpus with the lowest BPM is 'Mileage' from the album Six Summers with a 61.778 bpm, the highest is 'Middle' from the album SIRIUS with 199.928 bpm. 

The novelty funtion for 'Mileage' shows a continuous proportional distribution of peaks based on the 'energy-based novelty function based on Spotify’s loudness estimates'. In comparison to the function for the track 'Middle' we see that it contains a proportional distribution at the beginning but seems to half the distance of the peaks in a later section.This suggest that the track is around 200 BPM but is 'interpreted' in a BPM of 100 at the beginning which transforms in a interpretation in of the 200 BPM in the later section. 

```{r}
```

### NOVELTY FUNCTION MILEAGE

```{r}
library(tidyverse)
library(spotifyr)
library(compmus)

Mileage <-
  get_tidy_audio_analysis("4tOoKxtiLr1qH2AHUI5KqU") |>
  select(segments) |>
  unnest(segments)

Mileage |>
  mutate(loudness_max_time = start + loudness_max_time) |>
  arrange(loudness_max_time) |>
  mutate(delta_loudness = loudness_max - lag(loudness_max)) |>
  ggplot(aes(x = loudness_max_time, y = pmax(0, delta_loudness))) +
  geom_line() +
  xlim(0, 30) +
  theme_minimal() +
  labs(x = "Time (s)", y = "Novelty")

```

### NOVELTY FUNCTION MIDDLE

```{r}
library(tidyverse)
library(spotifyr)
library(compmus)

Middle <-
  get_tidy_audio_analysis("0JFkkqgv6r2SpbEAj7q5rt") |>
  select(segments) |>
  unnest(segments)

Middle |>
  mutate(loudness_max_time = start + loudness_max_time) |>
  arrange(loudness_max_time) |>
  mutate(delta_loudness = loudness_max - lag(loudness_max)) |>
  ggplot(aes(x = loudness_max_time, y = pmax(0, delta_loudness))) +
  geom_line() +
  xlim(0, 30) +
  theme_minimal() +
  labs(x = "Time (s)", y = "Novelty")

```

Internal relations {.storyboard}
=========================================

### INTERNAL RELATIONS
I made a 'bigger structure' dendogram based on all the individual tracks from the corpus to see how the tracks from all the different albums relate to each other. What strikes me at first when I look at the main structure and lines of the dendogram is that I see a clear distinction in classification between the different albums. There is a clear division per album visible. From this you could conclude that each album surrounds itself with songs within a specific genre of music and producers who produce that specific genre of music. This could also relate to the sound that is 'trending' at the moment an album is released. Maybe the producers produced the underlying music within this specific trend which creates a great similarity in sound of all the tracks combined in the specific album.

```{r}
```

### DENDOGRAM OF THE ENTIRE CORPUS

```{r}
library(tidyverse)
library(tidymodels)
library(ggdendro)
library(heatmaply)
library(spotifyr)
library(compmus)

get_conf_mat <- function(fit) {
  outcome <- .get_tune_outcome_names(fit)
  fit |> 
    collect_predictions() |> 
    conf_mat(truth = outcome, estimate = .pred_class)
}  

get_pr <- function(fit) {
  fit |> 
    conf_mat_resampled() |> 
    group_by(Prediction) |> mutate(precision = Freq / sum(Freq)) |> 
    group_by(Truth) |> mutate(recall = Freq / sum(Freq)) |> 
    ungroup() |> filter(Prediction == Truth) |> 
    select(class = Prediction, precision, recall)
}  

sevnalbums <-
  get_playlist_audio_features("jnk", "36UKFGHMewsc9n8KIspNrB") |>
  add_audio_analysis() |>
  mutate(
    segments = map2(segments, key, compmus_c_transpose),
    pitches =
      map(segments,
          compmus_summarise, pitches,
          method = "mean", norm = "manhattan"
      ),
    timbre =
      map(
        segments,
        compmus_summarise, timbre,
        method = "mean"
      )
  ) |>
  mutate(pitches = map(pitches, compmus_normalise, "clr")) |>
  mutate_at(vars(pitches, timbre), map, bind_rows) |>
  unnest(cols = c(pitches, timbre))

sevnalbums_juice <-
  recipe(
    track.name ~
      danceability +
      energy +
      loudness +
      speechiness +
      acousticness +
      instrumentalness +
      liveness +
      valence +
      tempo +
      duration +
      C + `C#|Db` + D + `D#|Eb` +
      E + `F` + `F#|Gb` + G +
      `G#|Ab` + A + `A#|Bb` + B +
      c01 + c02 + c03 + c04 + c05 + c06 +
      c07 + c08 + c09 + c10 + c11 + c12,
    data = sevnalbums
  ) |>
  step_center(all_predictors()) |>
  step_scale(all_predictors()) |> 
  # step_range(all_predictors()) |> 
  prep(sevnalbums |> mutate(track.name = str_trunc(track.name, 20))) |>
  juice() |>
  column_to_rownames("track.name")

sevnalbums_dist <- dist(sevnalbums_juice, method = "euclidean")

sevnalbums_dist |> 
  hclust(method = "complete") |> # Try single, average, and complete.
  dendro_data() |>
  ggdendrogram()

```

Conclusion {.storyboard}
=========================================

### CONCLUSION OF MY RESEARCH
After my research, I think I can formulate a concrete, short-tempered conclusion to the main research question: ‘’Which Sevn Alias albums were the most commercially successful and does this commercial success correspond in any way to the label where the artist was at the time signed at?’’

After my entire analysation, I think the most recent album is at the moment the most popular due to its recent release date, but I think that the older albums can be seen as equally as popular and with the most commercial success. I think the people see Sevn Alias as a style icon for the modern day Dutch Hip Hop and that hey have a lot more compassion for the older albums where these people 'grew up with' through the evolution of the genre in Holland. I think people contemplate these albums as pure 'classics' which will never die out. 

In relation to this, I think the label Rotterdam Airlines mainly focussed on more commercial tracks with a higher danceability, which in these time periods did result to the biggest hits from Sevn Alias, but the later labels like 1OAK Entertainment, even under venture of Top Notch, did evolve him to making less 'commercial' tracks, but did stimulate popularity for bringing him back to the basics. Something I think his fans appreciate the most.

```{r}
```
